Progress Report for CSS 586 Project: \\ Music Generation Using Deep Learning

### 2021-05-04 papers/progress/jack/progress-jack.tex
Update to Jack's report

\subsection{Datasets}
We use a dataset called MusicNet published by University of Washington, the department of Com[uter Science 
\cite{thickstun2017learning}. This dataset is a collection of 330 freely-licensed classical music recordings 
captured under various studio and microphone conditions. The recordings are in .wav raw audio format and are
polyphonic, meaning that multiple musical notes could present at a single point in time.

\subsection{Data Processing}

\subsection{Model Building}

\section{Progress}

\section{Future Work}

>>>>

\subsection{Datasets}
We use a dataset called MusicNet published by University of Washington, the department of Computer Science 
\cite{thickstun2017learning}. This dataset is a collection of 330 freely-licensed classical music recordings 
captured under various studio and microphone conditions. The recordings are in .wav raw audio format and are
polyphonic, meaning that multiple musical notes could present at a single time point. Furthermore, the music is
performed by multiple instruments, which have unique sound signatures.

\subsection{Data Processing}
Each training example is a raw sound fragment, which has a length of several seconds, taken from an audio recording.
We will transform the time-domain sample to the mel spectrogram representation. The mel scale is a logarithmic
transformation of a signal frequency. The idea of this transformation is to replicate the human perception of the 
difference between two sounds of different frequencies. For example, a person can easily tell the difference between 
100 Hz and 200 Hz, but cannot distinguish 1000 Hz and 1100 Hz. This transformation is done by splitting the sample
into overlapping windows, performing the Short Time Fourier Transformation on each window, converting the frequencies
to mel scale, and finally converting the amplitude to decibels. Librosa, a python package for music analysis, has
built-in functions for performing this tranformation. An example of a mel spectrogam is shown in Figure \ref{spectrogram}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{mel_spec.png}
  \caption{Mel spectrogram of a 10 second sound segment from the dataset. The color represents the amplitude in dB.}
  \label{spectrogram}
\end{figure}

\subsection{Model Building}
We will implement sequence models such as RNN, GRU, and LSTM to model the mel spectrogram and use these models to
generate novel spectrograms that could be converted back to audio.

\section{Progress}
Initially, a lot of time was spent on understanding different representation of the input data. We also learned 
about the basics of sequence models such as RNN, GRU, and LSTM. Because the dataset is large and cannot be loaded into 
the memory at once, we are working on a data pipeline that processes the raw audio files as described in the Section 
Data Processing and presents training examples one by one to a model.
